---
title: "China Pregnancy Cohort (Ng/Chen et al, 2021)"
author: "Sherrianne Ng"
date: "17/09/2021"
output:
  html_document:
    toc: true
    toc_depth: 6
---

***

The analysis codes and processed datasets pertain to the following publication: 

Ng/Chen, et al. The pregnancy vaginal microbiome, sialidase activity, leukocyte presence and lower risk of preterm birth in a Chinese population. Journal X,X, doi: (2021).


***

# Setup

```{r setup, message=FALSE, warning=FALSE}
# Load packages
library(broom)
library(caret)
library(cowplot)
library(corrplot)
library(ggpubr)
library(grid)
library(here)
library(patchwork)
library(performance)
library(PerformanceAnalytics)
library(psych)
library(reactable)
library(survey)
library(tidyverse)
library(viridis)

# Functions 
source(here("Functions", "RadialDonutPlot.R")) # Radial plot
source(here("Functions", "CorrelationPlot.R")) # Correlation plots
source(here("Functions", "MicrobiomeProfiles.R")) # Microbiome profiles 
source(here("Functions", "BarPlots.R")) # Bar plots
```

***

# Data

## Metadata 

This is supplementary data file 1 available with manuscript. 

```{r}
# Full Cohort (n=2796) 
fullcohort <- read.csv(here("Data", "NgChen_SupplementaryTable1.csv"), header=TRUE) %>%
  filter(SampleType == "Vaginal swab")

# Final Cohort (n=2646)
finalcohort <- fullcohort %>%
  filter(SampleTrimesters != "3rd") %>%
  filter(VMG %in% c("VMG I", "VMG II", "VMG III", "VMG IV-A", "VMG IV-B", "VMG V", 
                    "VMG VI", "VMG VII"))

# Outcome Cohort (n=1379)
outcomecohort <- filter(finalcohort, !is.na(Age) & !is.na(BirthGestation))

# Sub-cohort (n=819)
subcohort <- outcomecohort %>%
  filter(!is.na(SialidaseActivity) & !is.na(Leukocyte) & !is.na(Chorioamnionitis)) %>%
  mutate(LactoDomDep = factor(LactoDomDep, levels = c("Dominant", "Deplete")),
         SialidaseActivity = factor(SialidaseActivity, levels = c("High", "Low")),
         Leukocyte = factor(Leukocyte, levels = c("High", "Low")),
         Chorioamnionitis = factor(Chorioamnionitis, levels = c("Chorio+", "Chorio-")))
```

## Relative Abundance 

The relative abundance data for the final cohort (n=2646) based on counts summed by most dominant genera and species. 

```{r}
# Dominant genera
fullcohort_genera <- read.csv(here("Data", "FinalCohort-2646_DominantGenera_RelativeAbundance.csv"), header = TRUE)

# Dominant species
fullcohort_species <- read.csv(here("Data", "FinalCohort-2646_DominantSpecies_RelativeAbundance.csv"), header = TRUE)
```

***

# Main Figures 

## Figure 1a 

```{r fig.width=14, fig.height=22}
# Panel 1: Dominant Species Group
Fig1a_panel1 <- ggplot(finalcohort, aes(x = DomSpecGroup, fill = DomSpecGroup)) + 
  geom_bar() + 
  scale_fill_viridis(discrete = TRUE) + 
  geom_text(stat = 'count', aes(label = ..count..), 
            position = position_dodge(0.9), vjust = -0.2 , size = 5.5) +
  labs(title = "Dominant Species Groups") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, size = 26), 
        axis.title.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.title.y = element_text(size = 26), 
        axis.text.y = element_text(size = 24),
        legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white", colour = "black"))

# Panel 2: Relative Abundance
Fig1a_panel2 <- ggplot(finalcohort, 
                       aes(x = DomSpecGroup, 
                           y = DomSpecGroup_Proportion, 
                           fill = DomSpecGroup)) +
  stat_boxplot(geom = 'errorbar', width = 0.5) +
  geom_boxplot() + 
  scale_fill_viridis(discrete = TRUE) + 
  labs(y = "Relative Abundance") + 
  theme_bw() + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.title.y = element_text(size = 26), 
        axis.text.y = element_text(size = 24), 
        legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white", colour = "black"))

# Panel 3: Shannon Diversity Index
labels <- c("Aerococcus spp.", "Anaerococcus spp.", "Atopobium vaginae",  
            "Bifidobacterium spp.", "Gardnerella spp.", "Lactobacillus crispatus", 
            "Lactobacillus delbrueckii", "Lactobacillus gasseri", "Lactobacillus iners",
            "Lactobacillus jensenii", "Lactobacillus spp.", "Megasphaera spp.",
            "Prevotella spp.", "Shuttleworthia spp.", "Streptococcus agalactiae", 
            "Streptococcus anginosus", "Streptococcus luteciae", "Streptococcus spp.",
            "Veillonella spp.")  
Fig1a_panel3 <- ggplot(finalcohort, 
                       aes(x = DomSpecGroup, 
                           y = ShannonDiversity, 
                           fill = DomSpecGroup)) +
  stat_boxplot(geom = 'errorbar', width = 0.5) + 
  geom_boxplot() + 
  scale_fill_viridis(discrete = TRUE, guide = guide_legend(nrow = 5)) + 
  scale_x_discrete(labels = labels) +
  labs(x = "Dominant Species", 
       y = "Shannon Diversity Index", 
       fill = "Species") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 26), 
        axis.text.x = element_text(size = 22, face = "italic"), 
        axis.title.y = element_text(size = 26), 
        axis.text.y = element_text(size = 24), 
        legend.position = "none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white", colour = "black")) +
  guides(x = guide_axis(angle = 90))

# Output plot
# png(here("Output", "Figure1a.png"), width = 2400, height = 4600, res = 200)
plot_grid(Fig1a_panel1, Fig1a_panel2, Fig1a_panel3, ncol = 1, rel_heights = c(1,1,1.7),
          align = "v")
# dev.off()
```

***

## Figure 1b

```{r fig.width=15, fig.height=15}
# Merge metadata and species relative abundance for outcome cohort
outcomecohort_meta_species <- 
  merge(outcomecohort, fullcohort_species, 
        by = "Sample_ID", all = FALSE) %>%
  select(Sample_ID, VMG, Lactobacillus_crispatus:Veillonella_spp.) %>%
  arrange(VMG) %>%
  rowid_to_column(var = "RowNum") %>%
  pivot_longer(cols = Lactobacillus_crispatus:Veillonella_spp.) %>%
  rename(Species = "name") %>%
  mutate(VMG = factor(VMG,
                      levels = c("VMG I", "VMG II", "VMG III", "VMG IV-A", 
                                 "VMG IV-B", "VMG V", "VMG VI", "VMG VII")),
         Species = factor(Species, levels = species_levels))

# Aesthetics
species_color <- c("#1f78b4", "#33a02c", "#ff7f00", "#e31a1c", "#e31a1c", 
                   "#e31a1c", "#e31a1c", "#e31a1c", "#e31a1c", "#fb9a99", 
                   "#fb9a99", "#fb9a99", "#fb9a99", "#fb9a99", "#fb9a99",
                   "#6a3d9a", "#a6cee3", "#a6cee3", "#cab2d6", "#C0C0C0")
vmg_color <- c("#1f78b4", "#33a02c", "#ff7f00", "#e31a1c", 
               "#fb9a99", "#6a3d9a", "#a6cee3", "#cab2d6")

# Draw plot
radial <- myRadialStackedBarPlot(outcomecohort_meta_species)
donut <- myDonutPlot(outcomecohort_meta_species, "VMG (n=1379)")
Figure1b <- ggdraw() + 
  draw_plot(radial, 0, 0, 1, 1, 0.875) + 
  draw_plot(donut, 0.083, 0.094, 1, 1) + 
  theme(plot.margin = unit(c(1, 3, 1, 1), "cm"))

# Output plot
# png(here("Output", "Figure1b.png"), width = 3100, height = 3000, res = 200)
Figure1b
# dev.off()
```

***

## Figure 1c

```{r fig.width=15, fig.height=15}
# Merge metadata and species relative abundance for final cohort
finalcohort_meta_species <- merge(finalcohort, fullcohort_species, 
                                  by = "Sample_ID", all = FALSE) %>%
  select(Sample_ID, Lactobacillus_crispatus:Veillonella_spp.)
  
# Correlation data
cor_A <- cor(finalcohort_meta_species[,2:21], method = "spearman")
cor_B <- cor(finalcohort_meta_species[,2:21], method = "spearman")

# Matrix of the p-value of the correlation
p.mat <- cor.mtest(finalcohort_meta_species[,2:21])

# Correlation plot 
# png(here("Output", "Figure1c.png"), width = 3000, height = 3000, res = 200)
corrplot(cor_A, mar = c(1,0,1,0), type = "upper", addCoef.col = "black", tl.pos = "tl",
         hclust.method = "ward.D2", number.font = 1, number.cex = 1.3, 
         tl.srt = 45, tl.cex = 1.7, cl.cex = 1.5)
corrplot(cor_B, mar = c(1,0,1,0), type = "lower", tl.pos = "tl", add = T, 
         insig = "label_sig", p.mat = p.mat, sig.level = c(.001, .01, .05), 
         hclust.method = "ward.D2", tl.srt = 45, pch.cex = 2, tl.cex = 1.7, 
         cl.cex = 1.5)
# dev.off()
```

***

## Figure 2a

> Plot

```{r fig.width=14, fig.height=11}
# Merge metadata and genera relative abundance for sub-cohort
subcohort_meta_genera <- merge(subcohort, fullcohort_genera, 
                          by = "Sample_ID", all = FALSE) 

# Count
count(subcohort_meta_genera, SialidaseActivity)

# Prepare diversity and richness data for women with high and low sialidase
highsial <- func_sialdata(subcohort_meta_genera, sialidase = 'High')
lowsial <- func_sialdata(subcohort_meta_genera, sialidase = 'Low')

# Prepare microbiome profiles data for women with high and low sialidase 
highsial_mp <- func_sialdata_mp(subcohort_meta_genera, sialidase = 'High')
lowsial_mp <- func_sialdata_mp(subcohort_meta_genera, sialidase = 'Low')

# Panel 1: High Sialidase (Microbiome Profile)
Fig2a_panel1 <- 
  ggplot(highsial_mp, 
         aes(x = RowNumber, y = value, 
             fill = factor(Genus, levels = genus_label))) +
  scale_fill_manual(values = genus_color, guide = guide_legend(nrow = 3)) +
  geom_bar(stat = "identity", width = 1, preserve = "single") + 
  labs(title = "Microbiome Profiles (High Sialidase Activity Women), n=36", 
       y = "Relative Abundance", fill = "Genus") + 
  main_theme

# Panel 2: High Sialidase (Shannon Diversity)
Fig2a_panel2 <- 
  ggplot(highsial, 
         aes(x = RowNumber, y = IPSCohort_ShannonDiversity)) +
  geom_col(width = 1) + 
  labs(y = "Shannon") + 
  ylim(0,2) + 
  main_theme

# Panel 3: High Sialidase (Species Observed)
Fig2a_panel3 <- 
  ggplot(highsial, 
         aes(x = RowNumber, y = IPSCohort_SpeciesObserved)) + 
  geom_col(width = 1) + 
  ylim(0,40) + 
  labs(y = "S Obs") + 
  main_theme

# Panel 4: Low Sialidase (Microbiome Profiles)
Fig2a_panel4 <- 
  ggplot(lowsial_mp, 
         aes(x = RowNumber, y = value, 
             fill = factor(Genus, levels = genus_label))) +
  scale_fill_manual(values = genus_color, guide = guide_legend(nrow = 3)) +
  geom_bar(stat = "identity", width = 1, preserve = "single") + 
  labs(title = "Microbiome Profiles (Low Sialidase Activity Women), n=783",
       y = "Relative Abundance", fill = "Genus") + 
  main_theme

# Panel 5: Low Sialidase (Shannon Diversity)
Fig2a_panel5 <-
  ggplot(lowsial, 
         aes(x = RowNumber, y = IPSCohort_ShannonDiversity)) + 
  geom_col(width = 1) + 
  labs(y = "Shannon") + 
  main_theme

# Panel 6: Low Sialidase (Species Observed)
Fig2a_panel6 <- 
  ggplot(lowsial, 
         aes(x = RowNumber, y = IPSCohort_SpeciesObserved)) + 
  geom_col(width = 1) + 
  ylim(0,40) + 
  labs(y = "S Obs") + 
  main_theme

# Output plot
# png(here("Output", "Figure2a.png"), width = 2300, height = 2300, res = 200)
plot_grid(Fig2a_panel1, Fig2a_panel2, Fig2a_panel3, 
          Fig2a_panel4, Fig2a_panel5, Fig2a_panel6, 
          ncol = 1, rel_heights = c(6,2,2,6,2,2), 
          align = "v")
# dev.off()
```

> Legend 

```{r fig.width=13}
# Plot figure legend only
FigLegend <- 
  ggplot(lowsial_mp, 
         aes(x = RowNumber, y = value, 
             fill = factor(Genus, levels = genus_label))) +
  scale_fill_manual(values = genus_color, guide = guide_legend(nrow = 4)) +
  labs(fill = "Genus") + 
  geom_bar(stat = "identity", width = 1, preserve = "single") + 
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 14), 
        legend.position = "bottom")

# png(here("Output", "Figure2a_legend.png"), width = 2300, height = 220, res = 200)
legend <- get_legend(FigLegend)
as_ggplot(legend)
# dev.off()
```

***

## Figure 2b 

> Summary

```{r}
# Lactobacillus Abundance 
subcohort %>% 
  select(SialidaseActivity, LactoDomDep, VMG) %>%
  modify_if(is.character, as.factor) %>%
  split(.$SialidaseActivity) %>%
  map(summary, maxsum = 10)
```

> Statistics 

```{r}
# Fisher's Exact Test 
set.seed(1)
ft_res <- lapply(subcohort[,c(14,13)], 
                 fisher.test, y = subcohort$SialidaseActivity, 
                 simulate.p.value = TRUE, B = 2000)
print(ft_res)
```

> Plots

```{r fig.width=12, fig.height=5}
# Lactobacillus Abundance 
Figure2b_panel1 <- func_stackedbar_lactoabun(subcohort,
                                             x_var = "SialidaseActivity",
                                             lacto_colors = c("#1f78b4", "#e31a1c"),
                                             title = "Lactobacillus Abundance",
                                             subtitle = "p=2.403e-13",
                                             nrow_num = 2,
                                             legend_position = "bottom")


# VMG 
Figure2b_panel2 <- func_stackedbar_vmg(subcohort, 
                                       x_var = "SialidaseActivity",
                                       title = "Vaginal Microbiome Group",
                                       subtitle = "p=0.0005",
                                       nrow_num = 3,
                                       legend_position = "bottom")

# Combine plots 
Figure2b <- plot_grid(Figure2b_panel1, Figure2b_panel2, nrow = 1, align = "h", rel_widths = c(1,1.2))

# png(here("Output", "Figure2b.png"), width = 2300, height = 1200, res = 200)
Figure2b
# dev.off()
```

***

## Figure 2c

The figure was generated using LefSe tool on Huttenhower Galaxy (https://huttenhower.sph.harvard.edu/galaxy/).

**The input data used was:** `IPSCohort-819_HuttenhowerGalaxy_MASTER_SialidaseActivity.txt`

**The parameters used were as follows:**

**(A) Format Data for LefSe**

- Select whether the vectors (features and meta-data information) are listed in rows or columns: Rows
- Select which row to use as class: SialidaseActivity_Recode
- Select which row to use as subclass: no subclass
- Select which row to use as subject: Sample_ID
- Per-sample normalization of the sum of the values to 1M (recommended when very low values are present): Yes

**(B) LDA Effect Size (LEfSe)**

- Alpha value for the factorial Kruskal-Wallis test among classes: 0.05
- Alpha value for the pairwise Wilcoxon test between subclasses: 0.05
- Threshold on the logarithmic LDA score for discriminative features: 2.0
- Do you want the pairwise comparisons among subclasses to be performed only among the subclasses with the same name? No. 
- Set the strategy for multi-class analysis: All-against-all (more strict)

**(C) Plot LEfSe Results**

- Set text and label options (font size, abbreviations, ...): Advanced 
  - The title of the cladogram:	Not available.	
  - Number of label levels to be displayed (-1 means all):	1	
  - Maximum length of feature names:	60	
  - Title font size:	14	
  - Label font size:	12	
  - Class font size:	12
- Set some graphical options to personalize the output: Default
- Output format: png
- Set the dpi resolution of the output: 1200

**LefSe:** IPSCohort-819_SialidaseActivity_LefSe.png

```{r}
figure2c_lefse <- here("Output", "IPSCohort-819_SialidaseActivity_LefSe.png")
figdraw <- ggdraw() + draw_image(figure2c_lefse, scale = 1)
figdraw
```

**(D) Plot Cladogram**

- Set structural parameters of the cladogram: Default
- Set text and label options (font size, abbreviations, ...): Default
- Set some graphical options to personalize the output: Default
- Output format: png
- Set the dpi resolution of the output: 1200

**Cladogram:** IPSCohort-819_SialidaseActivity_Cladogram.png

```{r}
figure2c_cladogram <- here("Output", "IPSCohort-819_SialidaseActivity_Cladogram.png")
figdraw <- ggdraw() + draw_image(figure2c_cladogram, scale = 1)
figdraw
```

<style>
div.grey { background-color:#e3e6e9; border-radius: 5px; padding: 20px;}
</style>
<div class = "grey">

**NOTE** 

- Only the LefSe plot was used for Figure 2c.

</div>
<p>

***

## Figure 2d

> Statistics

The statistics shown here pertain only to *Lactobacillus*, *Gardnerella*, *Prevotella* and *Atopobium*. See Table S4 for all statistics. 

```{r}
# Merge metadata and species relative abundance for sub-cohort
subcohort_meta_genera <- merge(subcohort, fullcohort_genera, 
                          by = "Sample_ID", all = FALSE) 

# Mann-Whitney test
lapply(subcohort_meta_genera[c(23:24,27,26)], 
       function(x) wilcox.test(x ~ SialidaseActivity,
                               p.adjust.method = "bonferroni", 
                               data = subcohort_meta_genera))
```

> Plots 

```{r fig.width=8, fig.height=8}
# Aesthetics
sial_colors <- c("#c51b7d", "#4d9221")

# Select columns to plot
colNames <- names(subcohort_meta_genera)[c(23:24,27,26)]

# Boxplots 
plt_list <- list()

for(i in colNames){
  boxplot <- ggplot(subcohort_meta_genera, 
                    aes(x = SialidaseActivity,
                        y = .data[[i]], 
                        fill = SialidaseActivity)) +
  stat_boxplot(geom = 'errorbar', width = 0.5) + 
    geom_boxplot() + 
    scale_fill_manual(values = sial_colors) +
    labs(title = i, subtitle = "p<0.01", y = "Relative Abundance") + 
    geom_jitter(shape = 16, position = position_jitter(0.2)) + 
    theme_bw() + 
    theme(plot.title = element_text(hjust = 0.5, size = 19), 
          plot.subtitle = element_text(hjust = 0.5, size = 18),
          axis.title.x = element_blank(), 
          axis.text.x = element_text(size = 18), 
          axis.title.y = element_text(size = 19), 
          axis.text.y = element_text(size = 18), 
          legend.position = "none")
  
  plt_list[[i]] <- boxplot
}

# Combine plots
Figure2d <- wrap_plots(plt_list, ncol = 2)

# png(here("Output", "Figure2d.png"), width = 2000, height = 2000, res = 200)
Figure2d
# dev.off()
```

***

## Figure 2e

```{r}
# Counts 
count(subcohort, LactoDomDep)
count(subcohort, SialidaseActivity)

# View data
table(subcohort$SialidaseActivity, subcohort$LactoDomDep)

# Labels
TP <- 30
FP <- 6
FN <- 183
TN <- 600
sen <- TP+FN 
spec <- FP+TN 

# Confusion matrix
lvs <- c("no", "yes")
actual <- factor(rep(lvs, times = c(spec, sen)),
                levels = rev(lvs))
pred <- factor(
               c(
                 rep(lvs, times = c(TN, FP)),
                 rep(lvs, times = c(FN, TP))),
               levels = rev(lvs))

xtab <- table(pred, actual)

# Account for disease prevalence 15%
confusionMatrix(xtab, prevalence = 0.15)
```

***

## Figure 3a

> Summary

```{r}
subcohort %>% 
  select(LeukocyteGrades, LactoDomDep, VMG) %>% 
  modify_if(is.character, as.factor) %>%
  split(.$LeukocyteGrades) %>%
  map(summary)
```

> Statistics 

```{r}
# Fisher's Exact Test 
set.seed(1)
ft_res <- lapply(subcohort[,c(14,13)], 
                 fisher.test, y = subcohort$LeukocyteGrades, 
                 simulate.p.value = TRUE, B = 2000)
print(ft_res)
```

> Plots 

```{r fig.width=13, fig.height=4}
# Lactobacillus Abundance 
Figure3a_panel1_v2 <- 
  func_stackedbar_lactoabun(subcohort,
                            x_var = "LeukocyteGrades",
                            lacto_colors = c("#1f78b4", "#e31a1c"),
                            title = "Lactobacillus Abundance (p=0.003)", 
                            subtitle = "n=165        n=82        n=486          n=86",
                            nrow_num = 2,
                            legend_position = "right")

# VMG 
Figure3a_panel2_v2 <- 
  func_stackedbar_vmg(subcohort, 
                      x_var = "LeukocyteGrades",
                      title = "Vaginal Microbiome Group (p=0.02)",
                      subtitle = "n=165        n=82        n=486          n=86",
                      nrow_num = 8, 
                      legend_position = "right")

# png(here("Output", "Figure3a.png"), width = 3200, height = 1100, res = 200)
plot_grid(Figure3a_panel1_v2, Figure3a_panel2_v2, nrow = 1,
          align = "h")
# dev.off()
```

***

## Figure 3b 

> Statistics 

```{r}
# Merge metadata and species relative abundance for sub-cohort
subcohort_meta_species <- merge(subcohort, fullcohort_species, 
                          by = "Sample_ID", all = FALSE)

# Kruskall-Wallis test
set.seed(1)
lapply(subcohort_meta_species[c(23,26,24,27,32,31)],
       function(x) kruskal.test(x ~ LeukocyteGrades, 
                                data = subcohort_meta_species))

# Pairwise wilcox test 
set.seed(1)
pwt_res <- lapply(c(23,26,24,27,32,31), function(x) 
  pairwise.wilcox.test(subcohort_meta_species[[x]], 
                       subcohort_meta_species$LeukocyteGrades))
names(pwt_res) <- names(subcohort_meta_species)[c(23,26,24,27,32,31)]
pwt_res
```

> Plots 

```{r fig.width=18, fig.height=4}
# Aesthetics
leukgrade_colors <- c("#33a02c", "#33a02c", "#e31a1c", "#e31a1c")

# Rename columns 
subcohort_meta_species_x <- subcohort_meta_species %>%
  rename("L.crispatus" = "Lactobacillus_crispatus", 
         "L.gasseri" = "Lactobacillus_gasseri",
         "L.iners" = "Lactobacillus_iners", 
         "L.jensenii" = "Lactobacillus_jensenii",
         "L.delbrueckii" = "Lactobacillus_delbrueckii")

# Figure 3b
func_3b <- function(y_var, title, subtitle){
  # Boxplots
  ggplot(subcohort_meta_species_x, 
         aes(x = LeukocyteGrades,
             y = y_var,
             fill = LeukocyteGrades)) +
    stat_boxplot(geom = 'errorbar', width = 0.5) + 
    geom_boxplot() + 
    scale_fill_manual(values = leukgrade_colors) +
    labs(title = title, subtitle = subtitle, y = "Relative Abundance") + 
    geom_jitter(shape = 16, position = position_jitter(0.2)) + 
    theme_bw() + 
    theme(plot.title = element_text(hjust = 0.5, size = 19, face = "italic"), 
          plot.subtitle = element_text(hjust = 0.5, size = 18),
          axis.title.x = element_blank(), 
          axis.text.x = element_text(size = 18), 
          axis.title.y = element_text(size = 19), 
          axis.text.y = element_text(size = 18), 
          legend.position = "none")
}

Figure3b_p1 <- func_3b(subcohort_meta_species_x$L.crispatus, "L. crispatus", "p>0.05")
Figure3b_p2 <- func_3b(subcohort_meta_species_x$L.gasseri, "L. gasseri", "p>0.05")
Figure3b_p3 <- func_3b(subcohort_meta_species_x$L.iners, "L. iners", "p<0.01")
Figure3b_p4 <- func_3b(subcohort_meta_species_x$L.jensenii, "L. jensenii", "p>0.05")
Figure3b_p5 <- func_3b(subcohort_meta_species_x$L.delbrueckii, "L. delbrueckii", "p>0.05")
Figure3b_p6 <- func_3b(subcohort_meta_species_x$Lactobacillus_spp., "Lactobacillus spp.", "p>0.05")

# Combined figure
Figure3b <- plot_grid(Figure3b_p1, Figure3b_p2, Figure3b_p3, 
                      Figure3b_p4, Figure3b_p5, Figure3b_p6,
                      nrow = 1)

# png(here("Output", "Figure3b.png"), width = 2500, height = 550, res = 150)
Figure3b 
# dev.off()
```

***

## Figure 3c 

The figure was generated using LefSe tool on Huttenhower Galaxy (https://huttenhower.sph.harvard.edu/galaxy/).

**The input data used was:** `HuttenhowerGalaxy_Subcohort-823_Leukocyte.txt`

**The parameters used were as follows:**

**(A) Format Data for LefSe**

- Select whether the vectors (features and meta-data information) are listed in rows or columns: Rows
- Select which row to use as class: Leukocyte_Recode
- Select which row to use as subclass: no subclass
- Select which row to use as subject: Sample_ID
- Per-sample normalization of the sum of the values to 1M (recommended when very low values are present): Yes

**(B) LDA Effect Size (LEfSe)**

- Alpha value for the factorial Kruskal-Wallis test among classes: 0.05
- Alpha value for the pairwise Wilcoxon test between subclasses: 0.05
- Threshold on the logarithmic LDA score for discriminative features: 2.0
- Do you want the pairwise comparisons among subclasses to be performed only among the subclasses with the same name? No. 
- Set the strategy for multi-class analysis: All-against-all (more strict)

**(C) Plot LEfSe Results**

- Set text and label options (font size, abbreviations, ...): Advanced
  - The title of the cladogram	Not available.	
  - Number of label levels to be displayed (-1 means all)	1	
  - Maximum length of feature names	60	
  - Title font size	14	
  - Label font size	12	
  - Class font size	12
- Set some graphical options to personalize the output: Default
- Output format: png
- Set the dpi resolution of the output: 1200

**LefSe:** IPSCohort-819_Leukocyte_LefSe.png

```{r}
figure3c_lefse <- here("Output", "IPSCohort-819_Leukocyte_LefSe.png")
figdraw <- ggdraw() + draw_image(figure3c_lefse, scale = 1)
figdraw
```

**(D) Plot Cladogram**

- Set structural parameters of the cladogram: Default
- Set text and label options (font size, abbreviations, ...): Default
- Set some graphical options to personalize the output: Default
- Output format: png
- Set the dpi resolution of the output: 1200

**Cladogram:** IPSCohort-819_Leukocyte_Cladogram.png

```{r}
figure3c_cladogram <- here("Output", "IPSCohort-819_Leukocyte_Cladogram.png")
figdraw <- ggdraw() + draw_image(figure3c_cladogram, scale = 1)
figdraw
```

***

## Figure 4a

> Summary 

```{r}
# Outcome cohort 
outcomecohort %>% 
  select(Group_PretermTerm, LactoDomDep, VMG) %>%
  modify_if(is.character, as.factor) %>%
  split(.$Group_PretermTerm) %>%
  map(summary)
  
# Subcohort 
subcohort %>%
  select(Group_PretermTerm, SialidaseActivity, Leukocyte) %>%
  modify_if(is.character, as.factor) %>%
  split(.$Group_PretermTerm) %>%
  map(summary)
```

> Statistics 

```{r}
# Fisher's Exact Test: Outcome Cohort
set.seed(1)
ft_res <- lapply(outcomecohort[,c(14,13)], 
                 fisher.test, y = outcomecohort$Group_PretermTerm, 
                 simulate.p.value = TRUE, B = 2000)
print(ft_res)

# Fisher's Exact Test: Subcohort
set.seed(1)
ft_res <- lapply(subcohort[,c(16,18)], 
                 fisher.test, y = subcohort$Group_PretermTerm, 
                 simulate.p.value = TRUE, B = 2000)
print(ft_res)
```

> Plots

```{r fig.width=15, fig.height=10}
# Count preterm/term in outcome cohort and subcohort 
count(outcomecohort, Group_PretermTerm)
count(subcohort, Group_PretermTerm)

# Lactobacillus Abundance 
Figure4a_p1 <- 
  func_stackedbar_lactoabun(outcomecohort,
                            x_var = "Group_PretermTerm",
                            lacto_colors = c("#e31a1c", "#1f78b4"),
                            title = "Lactobacillus Abundance (p=0.57)",
                            subtitle = "n=72                            n=1307",
                            nrow_num = 1,
                            legend_position = "bottom")

# VMG 
Figure4a_p2 <- 
  func_stackedbar_vmg(outcomecohort, 
                      x_var = "Group_PretermTerm",
                      title = "Vaginal Microbiome Group (p=0.05)",
                      subtitle = "n=72                                n=1307",
                      nrow_num = 2,
                      legend_position = "bottom")

# Sialidase Activity 
Figure4a_p3 <- 
  func_stackedbar_sial(subcohort,
                       x_var = "Group_PretermTerm",
                       title = "Sialidase Activity (p=0.45)",
                       subtitle = "n=46                            n=773",
                       nrow_num = 1,
                       legend_position = "bottom")

# Leukocyte
Figure4a_p4 <- 
  func_stackedbar_leuk(subcohort,
                          x_var = "Group_PretermTerm",
                          title = "Leukocyte (p=0.74)",
                          subtitle = "n=46                                  n=773",
                          nrow_num = 1,
                          legend_position = "bottom")

# Combine plots 
Figure4a <- plot_grid(Figure4a_p1, Figure4a_p2, 
                      Figure4a_p3, Figure4a_p4,
                      nrow = 2, align = "h", rel_widths = c(1,1.2))

# png(here("Output", "Figure4a.png"), width = 3000, height = 2000, res = 200)
Figure4a
# dev.off()
```

***

## Figure 4b

> Statistics 

```{r}
fig4b_stat <- subcohort %>%
  mutate(StatGroup = case_when(
    LactoDomDep == "Deplete" & SialidaseActivity == "High" & Leukocyte == "High" ~ "A",
    LactoDomDep == "Dominant" & SialidaseActivity == "High" & Leukocyte == "High" ~ "B",
    LactoDomDep == "Deplete" & SialidaseActivity == "Low" & Leukocyte == "High" ~ "C",
    LactoDomDep == "Dominant" & SialidaseActivity == "Low" & Leukocyte == "High" ~ "D",
    LactoDomDep == "Deplete" & SialidaseActivity == "High" & Leukocyte == "Low" ~ "E",
    # LactoDomDep == "Dominant" & SialidaseActivity == "High" & Leukocyte == "Low" ~ "F",
    LactoDomDep == "Deplete" & SialidaseActivity == "Low" & Leukocyte == "Low" ~ "G",
    LactoDomDep == "Dominant" & SialidaseActivity == "Low" & Leukocyte == "Low" ~ "H"))

# Kruskal-Wallis
kruskal.test(BirthGestation ~ StatGroup, data = fig4b_stat)
```

> Plots 

```{r fig.width=10, fig.height=10}
# Convert gestation in days to weeks
subcohort_weeks <- subcohort %>%
  mutate(BirthGestation_Weeks = BirthGestation/7)

# Aesthetics
vmg_color <- c("#1f78b4", "#33a02c", "#ff7f00", "#e31a1c",
                  "#fb9a99", "#6a3d9a", "#a6cee3", "#cab2d6")  
ldd_colors <- c("#e31a1c", "#1f78b4")

# Facet labels
sial.labs <- c("High Sialidase Activity", "Low Sialidase Activity")
names(sial.labs) <- c("High", "Low")
leuk.labs <- c("Leukocyte High", "Leukocyte Low")
names(leuk.labs) <- c("High", "Low")
  
# Plot 
Figure4b <- ggplot(subcohort_weeks,
                   aes(x = LactoDomDep,
                       y = BirthGestation_Weeks,
                       fill = LactoDomDep)) +
  stat_boxplot(geom = 'errorbar', width = 0.5) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), 
              size = 2.5,
              alpha = 0.9,
              aes(colour = VMG)) + 
  scale_colour_manual(values = vmg_color) +
  scale_fill_manual(values = ldd_colors, guide = "none") +
  facet_grid(Leukocyte ~ SialidaseActivity,
             labeller = labeller(SialidaseActivity = sial.labs,
                                 Leukocyte = leuk.labs)) +
  labs(title = "Vaginal Microbiome, Sialidase Activity, Leukocytes and Birth Outcome", 
       subtitle = "n=819",
       x = "Lactobacillus Abundance",
       y = "Birth Gestation (Weeks)") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, size = 19),
        plot.subtitle = element_text(hjust = 0.5, size = 18),
        axis.title.x = element_text(size = 19),
        axis.text.x = element_text(size = 18),
        axis.title.y = element_text(size = 19),
        axis.text.y = element_text(size = 18),
        strip.background = element_rect(color = "black", size = 1, linetype = "solid"), 
        strip.text.x = element_text(size = 18, color = "black"),
        strip.text.y = element_text(size = 18, color = "black"),
        legend.title = element_text(size = 19),
        legend.text = element_text(size = 18),
        legend.position = "bottom")
  
# png(here("Output", "Figure4b.png"), width = 2000, height = 2000, res = 200)
Figure4b
# dev.off()
```

***

# Supplementary Figures 

## Figure S1a 

```{r}
# Full Cohort 
count(fullcohort)

# Minimum library size 
fullcohort %>% 
  mutate(MinLibSize = case_when(
    is.na(DomSpecGroup) ~ "Exclude",
    !is.na(DomSpecGroup) ~ "Include")) %>%
  count(MinLibSize)

# Microbiome Classifications 
fullcohort %>% 
  filter(!is.na(DomSpecGroup)) %>%
  count(VMG, DomSpecGroup)

# Full Cohort: Trimesters 
count(fullcohort, SampleTrimesters)

# Final Cohort: Women with and without outcome
finalcohort %>%
  mutate(OutcomeGroup = case_when(
    !is.na(Age) & !is.na(BirthGestation) ~ "Outcome",
    is.na(Age) | is.na(BirthGestation) ~ "NoOutcome")) %>%
  count(OutcomeGroup)

# Sub-cohort 
count(subcohort)
```

***

## Figure S1b 

```{r fig.width=10}
# Annotate women with and without outcome data then assign levels
finalcohort_group <- finalcohort %>%
  filter(SampleType == "Vaginal swab") %>%
  mutate(OutcomeGroup = case_when(!is.na(Age) & !is.na(BirthGestation) ~ "Outcome", 
                                  is.na(Age) | is.na(BirthGestation) ~ "NoOutcome"),
         OutcomeGroup = factor(OutcomeGroup,
                               levels = c("Outcome", "NoOutcome")), 
         SampleGestation_Weeks = SampleGestation/7)

# Count
count(finalcohort_group, OutcomeGroup)

# Aesthetics
outcome_colors <- c("#e31a1c", "#1f78b4")

# Plot density plot
FigureS1b <- 
  ggplot(finalcohort_group, 
         aes(x = SampleGestation_Weeks, fill = OutcomeGroup)) +
  geom_histogram(aes(y = ..density..), 
                 alpha = 0.8, 
                 position = "identity", 
                 binwidth = 0.2) +
  geom_density(alpha = .4) + 
  scale_x_continuous(breaks = seq(10,40,5)) +
  scale_fill_manual(values = outcome_colors) + 
  labs(title = "Sample Gestations",
       subtitle = "Outcome (n=1379), No Outcome (n=1267)",
       x = "Sample Gestation (Weeks)",
       fill = "OutcomeGroup") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, size = 19),
        plot.subtitle = element_text(hjust = 0.5, size = 18),
        axis.title.x = element_text(size = 19), 
        axis.text.x = element_text(size = 18), 
        axis.title.y = element_text(size = 19), 
        axis.text.y = element_text(size = 18), 
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 15),
        legend.position = "bottom")


# png(here("Output", "FigureS1b.png"), width = 2000, height = 1200, res = 200)
FigureS1b
# dev.off()
```

***

## Figure S2

```{r fig.width=13, fig.height=14}
# Labels
labels <- c("Aerococcus spp.", "Anaerococcus spp.", "Atopobium vaginae",  
            "Bifidobacterium spp.", "Gardnerella spp.", "Lactobacillus crispatus", 
            "Lactobacillus delbrueckii", "Lactobacillus gasseri", "Lactobacillus iners",
            "Lactobacillus jensenii", "Lactobacillus spp.", "Megasphaera spp.",
            "Prevotella spp.", "Shuttleworthia spp.", "Streptococcus agalactiae", 
            "Streptococcus anginosus", "Streptococcus luteciae", "Streptococcus spp.",
            "Veillonella spp.")  

# Shannon Diversity: non-normalised data
FigS2_p1 <- ggplot(finalcohort, 
                   aes(x = DomSpecGroup,
                       y = ShannonDiversity,
                       fill = DomSpecGroup)) +
  stat_boxplot(geom = 'errorbar', width = 0.5) + 
  geom_boxplot() + 
  scale_fill_viridis(discrete = TRUE, guide = guide_legend(nrow = 5)) + 
  scale_x_discrete(labels = labels) +
  labs(x = "Dominant Species", 
       y = "Shannon Diversity Index", 
       fill = "Species") + 
  theme_bw() +
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(), 
        axis.title.y = element_text(size = 26), 
        axis.text.y = element_text(size = 24), 
        legend.position = "none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white", colour = "black")) +
  guides(x = guide_axis(angle = 90))

# Shannon Diversity: TMM normalised data
FigS2_p2 <- ggplot(finalcohort,
                   aes(x = DomSpecGroup,
                       y = ShannonDiversity.tmm,
                       fill = DomSpecGroup)) +
  stat_boxplot(geom = 'errorbar', width = 0.5) + 
  geom_boxplot() + 
  scale_fill_viridis(discrete = TRUE, guide = guide_legend(nrow = 5)) + 
  scale_x_discrete(labels = labels) +
  labs(x = "Dominant Species", 
       y = "Shannon Diversity Index (TMM)", 
       fill = "Species") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 26), 
        axis.text.x = element_text(size = 22, face = "italic"), 
        axis.title.y = element_text(size = 26), 
        axis.text.y = element_text(size = 24), 
        legend.position = "none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white", colour = "black")) +
  guides(x = guide_axis(angle = 90))

# Output plot
# png(here("Output", "FigureS2.png"), width = 2400, height = 2900, res = 200)
plot_grid(FigS2_p1, FigS2_p2, ncol = 1, align = "v", rel_heights = c(1,1.7))
# dev.off()
```

***

## Figure S3 

```{r fig.width=15, fig.height=15}
# Extract no outcome cohort 
nooutcomecohort <- finalcohort %>%
  filter(is.na(Age) | is.na(BirthGestation))

# Merge metadata and species relative abundance for no outcome cohort
nooutcomecohort_meta_species <- 
  merge(nooutcomecohort, fullcohort_species, 
        by = "Sample_ID", all = FALSE) %>%
  select(Sample_ID, VMG, Lactobacillus_crispatus:Veillonella_spp.) %>%
  arrange(VMG) %>%
  rowid_to_column(var = "RowNum") %>%
  pivot_longer(cols = Lactobacillus_crispatus:Veillonella_spp.) %>%
  rename(Species = "name") %>%
  mutate(VMG = factor(VMG,
                      levels = c("VMG I", "VMG II", "VMG III", "VMG IV-A", 
                                 "VMG IV-B", "VMG V", "VMG VI", "VMG VII")),
         Species = factor(Species, levels = species_levels))

# Aesthetics
species_color <- c("#1f78b4", "#33a02c", "#ff7f00", "#e31a1c", "#e31a1c", 
                   "#e31a1c", "#e31a1c", "#e31a1c", "#e31a1c", "#fb9a99", 
                   "#fb9a99", "#fb9a99", "#fb9a99", "#fb9a99", "#fb9a99",
                   "#6a3d9a", "#a6cee3", "#a6cee3", "#cab2d6", "#C0C0C0")
vmg_color <- c("#1f78b4", "#33a02c", "#ff7f00", "#e31a1c", 
               "#fb9a99", "#6a3d9a", "#a6cee3", "#cab2d6")

# Draw plot
radial <- myRadialStackedBarPlot(nooutcomecohort_meta_species)
donut <- myDonutPlot(nooutcomecohort_meta_species, "VMG (n=1267)")
FigureS3 <- ggdraw() + 
  draw_plot(radial, 0, 0, 1, 1, 0.875) + 
  draw_plot(donut, 0.083, 0.094, 1, 1) + 
  theme(plot.margin = unit(c(1, 3, 1, 1), "cm"))

# Output plot
# png(here("Output", "FigureS3.png"), width = 3100, height = 3000, res = 200)
FigureS3
# dev.off()
```

***

## Figure S4

> Statistics

```{r}
# Fisher's Exact Test: Final Cohort (Lactobacillus abundance and VMG)
set.seed(1)
ft_res <- lapply(finalcohort[,c(14,13)], 
                 fisher.test, y = finalcohort$DNA_Extraction_Kit, 
                 simulate.p.value = TRUE, B = 2000)
print(ft_res)

# Fisher's Exact Test: Outcome Cohort (Birth Outcome)
set.seed(1)
fisher.test(outcomecohort$DNA_Extraction_Kit, outcomecohort$Group_PretermTerm, 
            simulate.p.value = TRUE, B = 2000)

# Fisher's Exact Test: Subcohort (Chorioamnionitis)
set.seed(1)
fisher.test(subcohort$DNA_Extraction_Kit, subcohort$Chorioamnionitis, 
            simulate.p.value = TRUE, B = 2000)
```

> Summary 

```{r}
# Lactobacillus abundance
finalcohort %>% 
  group_by(DNA_Extraction_Kit) %>% 
  count(LactoDomDep) %>% 
  mutate(Percent = round(n/sum(n)*100, 2),
         Percentage = round(n/sum(n)*100)) %>% 
  ungroup() %>%
  as.data.frame()

# VMG
finalcohort %>% 
  group_by(DNA_Extraction_Kit) %>% 
  count(VMG) %>% 
  mutate(Percent = round(n/sum(n)*100, 2),
         Percentage = round(n/sum(n)*100)) %>% 
  ungroup() %>%
  as.data.frame()

# Birth outcome 
outcomecohort %>% 
  group_by(DNA_Extraction_Kit) %>% 
  count(Group_PretermTerm) %>%
  mutate(Percent = round(n/sum(n)*100, 2),
         Percentage = round(n/sum(n)*100)) %>% 
  ungroup() %>%
  as.data.frame()

# Chorioamnionitis
subcohort %>%
  group_by(DNA_Extraction_Kit) %>%
  count(Chorioamnionitis) %>%
  mutate(Percent = round(n/sum(n)*100, 2),
         Percentage = round(n/sum(n)*100)) %>% 
  ungroup() %>%
  as.data.frame()
```

> Plots 

```{r fig.width=15, fig.height=10}
# Lactobacillus abundance 
count(finalcohort, DNA_Extraction_Kit)
FigS4_p1 <- func_stackedbar_lactoabun(
  finalcohort,
  x_var = "DNA_Extraction_Kit",
  lacto_colors = c("#e31a1c", "#1f78b4"),
  title = "Lactobacillus Abundance (p=0.96)",
  subtitle = "n=1510                                 n=1136",
  nrow_num = 1,
  legend_position = "bottom")

# VMG 
FigS4_p2 <- func_stackedbar_vmg(
  finalcohort, 
  x_var = "DNA_Extraction_Kit",
  title = "Vaginal Microbiome Group (p=0.83)",
  subtitle = "n=1510                                 n=1136",
  nrow_num = 2,
  legend_position = "bottom")

# Preterm/Term
count(outcomecohort, DNA_Extraction_Kit)
FigS4_p3 <- func_stackedbar_ptb(
  outcomecohort, 
  x_var = "DNA_Extraction_Kit", 
  title = "Birth Outcome (p=0.47)", 
  subtitle = "n=764                               n=615", 
  nrow_num = 1, 
  legend_position = "bottom")

# Chorioamnionitis
count(subcohort, DNA_Extraction_Kit)
FigS4_p4 <- func_stackedbar_chorio(
  subcohort, 
  x_var = "DNA_Extraction_Kit", 
  title = "Chorioamnionitis (p=0.06)", 
  subtitle = "n=380                               n=439", 
  nrow_num = 1, 
  legend_position = "bottom")

# png(here("Output", "FigureS4.png"), width = 3100, height = 2500, res = 200)
plot_grid(FigS4_p1, FigS4_p2, FigS4_p3, FigS4_p4,
          nrow = 2, align = "h")
# dev.off()
```

***

## Figure S5

```{r fig.width=15, fig.height=11}
# Merge metadata and species relative abundance for sub-cohort
subcohort_metaspec <- merge(subcohort, fullcohort_species, 
                          by = "Sample_ID", all = FALSE, sort = FALSE) %>%
  filter(LactoDomDep == "Dominant") %>%
  filter(VMG %in% c("VMG I", "VMG II", "VMG III", "VMG V")) %>%
  select(Sample_ID, VMG, SialidaseActivity, Lactobacillus_crispatus:Veillonella_spp.) %>%
  pivot_longer(cols = Lactobacillus_crispatus:Veillonella_spp.) %>%
  rename(Species = "name", Abundance = "value")

# Count 
subcohort_metaspec %>%
  distinct(Sample_ID, .keep_all = TRUE) %>%
  count(SialidaseActivity)

# High sialidase activity microbiome profiles
highsial_dom <- subcohort_metaspec %>% 
  filter(SialidaseActivity == "High") 
  
FigureS5_p1 <- 
  func_lacto_profiles(highsial_dom, 
                      title = "Species-level Microbiome Profiles", 
                      subtitle = "High Sialidase Activity + Lactobacillus Dominant (n=6)",
                      legend_position = "none",
                      nrow_num = 5)

# Low sialidase activity microbiome profiles
lowsial_dom <- subcohort_metaspec %>% 
  filter(SialidaseActivity == "Low") 
  
FigureS5_p2 <- 
  func_lacto_profiles(lowsial_dom, 
                      title = "", 
                      subtitle = "Low Sialidase Activity + Lactobacillus Dominant (n=597)",
                      legend_position = "bottom",
                      nrow_num = 5)


# png(here("Output", "FigureS5.png"), width = 3300, height = 2400, res = 200)
plot_grid(FigureS5_p1, FigureS5_p2, ncol = 1, rel_heights = c(1,1.22))
# dev.off()
```

***

## Figure S6

```{r}
# Assign PTB levels 
outcomecohort_ptb <- outcomecohort %>% 
  mutate(BirthGestation_weeks = BirthGestation/7,
         Group_Gestation = case_when(
           BirthGestation_weeks <34 ~ "EarlyPreterm",
           BirthGestation_weeks >=34 & BirthGestation_weeks <37 ~ "LatePreterm",
           BirthGestation_weeks >=37 ~ "Term"))

# Count groups 
count(outcomecohort_ptb, Group_Gestation)
```

> Statistics 

```{r}
# Fisher's Exact Test: Group Gestation - Lactobacillus Abundance
set.seed(1)
fisher.test(outcomecohort_ptb$Group_Gestation, outcomecohort_ptb$LactoDomDep, 
            simulate.p.value = TRUE, B = 2000)

# Fisher's Exact Test: Group Gestation - VMG
set.seed(1)
fisher.test(outcomecohort_ptb$Group_Gestation, outcomecohort_ptb$VMG, 
            simulate.p.value = TRUE, B = 2000)
```


> Plots 

```{r fig.width=15, fig.height=8}
# Group Gestation: Lactobacillus Abundance
FigureS6_p1 <- 
  func_stackedbar_lactoabun(outcomecohort_ptb,
                            x_var = "Group_Gestation",
                            lacto_colors = c("#e31a1c", "#1f78b4"),
                            title = "Lactobacillus Abundance (p=0.64)",
                            subtitle = "n=21                       n=51                   n=1307",
                            nrow_num = 1,
                            legend_position = "bottom")


# Group Gestation: VMG 
FigureS6_p2 <- 
  func_stackedbar_vmg(outcomecohort_ptb, 
                      x_var = "Group_Gestation",
                      title = "Vaginal Microbiome Group (p=0.13)",
                      subtitle = "n=21                       n=51                   n=1307",
                      nrow_num = 2,
                      legend_position = "bottom")

# png(here("Output", "FigureS6.png"), width = 3200, height = 1200, res = 200)
plot_grid(FigureS6_p1, FigureS6_p2, 
          ncol = 2, align = "h")
# dev.off()
```

***

## Figure S7

Ensure `BarPlots.R` script is loaded, see Figure 2b.

> Statistics

```{r}
# Fisher's Exact Test 
set.seed(1)
ft_res <- lapply(subcohort[,c(13,12,15,17)], 
                 fisher.test, y = subcohort$Chorioamnionitis, 
                 simulate.p.value = TRUE, B = 2000)
print(ft_res)
```

> Plots

```{r fig.width=15, fig.height=10}
# Prepare data for subcohort
subcohort_chorio <- subcohort %>%
  mutate(Chorioamnionitis = factor(Chorioamnionitis,
                                   levels = c("Chorio+", "Chorio-")),
         LactoDomDep = factor(LactoDomDep,
                              levels = c("Deplete", "Dominant")),
         SialidaseActivity = factor(SialidaseActivity,
                            levels = c("High", "Low")),
         Leukocyte = factor(Leukocyte,
                            levels = c("High", "Low")))

# Count number in each chorioamnionitis group
count(subcohort_chorio, Chorioamnionitis)

# Lactobacillus Abundance 
FigureS7b_p1 <- 
  func_stackedbar_lactoabun(subcohort_chorio,
                            x_var = "Chorioamnionitis",
                            lacto_colors = c("#e31a1c", "#1f78b4"),
                            title = "Lactobacillus Abundance (p=0.78)",
                            subtitle = "n=216                                 n=603",
                            nrow_num = 1,
                            legend_position = "bottom")


# VMG 
FigureS7b_p2 <- 
  func_stackedbar_vmg(subcohort_chorio, 
                      x_var = "Chorioamnionitis",
                      title = "Vaginal Microbiome Group (p=0.80)",
                      subtitle = "n=216                                 n=603",
                      nrow_num = 2,
                      legend_position = "bottom")

# Sialidase Activity 
FigureS7b_p3 <- 
  func_stackedbar_sial(subcohort_chorio,
                       x_var = "Chorioamnionitis",
                       title = "Sialidase Activity (p=0.44)",
                       subtitle = "n=216                                 n=603",
                       nrow_num = 1,
                       legend_position = "bottom")

# Leukocyte
FigureS7b_p4 <- 
  func_stackedbar_leuk(subcohort_chorio,
                       x_var = "Chorioamnionitis",
                       title = "Leukocyte (p=0.26)",
                       subtitle = "n=216                                 n=603",
                       nrow_num = 1,
                       legend_position = "bottom")


# png(here("Output", "FigureS7b.png"), width = 3100, height = 2500, res = 200)
plot_grid(FigureS7b_p1, FigureS7b_p2,
          FigureS7b_p3, FigureS7b_p4,
          nrow = 2, align = "h")
# dev.off()
```

***

# Supplementary Tables 

## Table S1

See NgChen_SupplementaryTable1.csv. 

## Table S2

```{r}
finalcohort %>% 
  select(DomSpecGroup, DomSpecGroup_Proportion, ShannonDiversity, ShannonDiversity.tmm) %>%
  split(.$DomSpecGroup) %>%
  map(summary)
```

***

## Table S3

```{r}
# Annotate women with and without outcome data 
finalcohort_group <- finalcohort %>%
  filter(SampleType == "TestSample") %>%
  mutate(OutcomeGroup = case_when(!is.na(Age) & !is.na(BirthGestation) ~ "Outcome", 
                                  is.na(Age) | is.na(BirthGestation) ~ "NoOutcome"),
         OutcomeGroup = factor(OutcomeGroup,
                               levels = c("Outcome", "NoOutcome")))

# Count women with and without outcome data
count(finalcohort_group, OutcomeGroup)

# Women with outcome data 
finalcohort_group %>%
  filter(OutcomeGroup == "Outcome") %>%
  count(VMG) %>%
  mutate(Percentage = round(n/sum(n)*100, 2)) %>%
  reactable()

# Women without outcome data 
finalcohort_group %>%
  filter(OutcomeGroup == "NoOutcome") %>%
  count(VMG) %>%
  mutate(Percentage = round(n/sum(n)*100, 2)) %>%
  reactable()
```

***

## Table S4

> Summary 

```{r}
# Merge metadata and species relative abundance for sub-cohort
subcohort_meta_genera <- merge(subcohort, fullcohort_genera, 
                          by = "Sample_ID", all = FALSE) %>%
  select(Sample_ID, SialidaseActivity, Atopobium, Bifidobacterium,
         Clostridium, Gardnerella, Lactobacillus, Megasphaera,
         Prevotella)

# Relative abundance of each genus by sialidase activity 
subcohort_meta_genera %>%
  split(.$SialidaseActivity) %>%
  map(summary)
```

> Statistics

```{r}
# Mann-Whitney Test: Atopobium
wilcox.test(Atopobium ~ SialidaseActivity, 
            subcohort_meta_genera, 
            p.adjust.method = "bonferroni")

# Mann-Whitney Test: Bifidobacterium
wilcox.test(Bifidobacterium ~ SialidaseActivity, 
            subcohort_meta_genera, 
            p.adjust.method = "bonferroni")

# Mann-Whitney Test: Clostridium
wilcox.test(Clostridium ~ SialidaseActivity, 
            subcohort_meta_genera, 
            p.adjust.method = "bonferroni")

# Mann-Whitney Test: Gardnerella
wilcox.test(Gardnerella ~ SialidaseActivity, 
            subcohort_meta_genera, 
            p.adjust.method = "bonferroni")

# Mann-Whitney Test: Lactobacillus
wilcox.test(Lactobacillus ~ SialidaseActivity, 
            subcohort_meta_genera, 
            p.adjust.method = "bonferroni")

# Mann-Whitney Test: Megasphaera
wilcox.test(Megasphaera ~ SialidaseActivity,
            subcohort_meta_genera, 
            p.adjust.method = "bonferroni")

# Mann-Whitney Test: Prevotella
wilcox.test(Prevotella ~ SialidaseActivity,
            subcohort_meta_genera, 
            p.adjust.method = "bonferroni")
         
```

***

## Table S5

The Stat Groups were as follows: 

**Group** | **Lactobacillus** | **Sialidase** | **Leukocyte**
:-----|:---------|:---------|:---------
A | Deplete | High | High 
B | Dominant | High | High 
C | Deplete | Low | High 
D | Dominant | Low | High
E | Deplete | High | Low 
F | Dominant | High | Low
G | Deplete | Low | Low
H | Dominant | Low | Low

```{r}
# Summary
subcohort %>%
  mutate(StatGroup = case_when(
    LactoDomDep == "Deplete" & SialidaseActivity == "High" & Leukocyte == "High" ~ "A",
    LactoDomDep == "Dominant" & SialidaseActivity == "High" & Leukocyte == "High" ~ "B",
    LactoDomDep == "Deplete" & SialidaseActivity == "Low" & Leukocyte == "High" ~ "C",
    LactoDomDep == "Dominant" & SialidaseActivity == "Low" & Leukocyte == "High" ~ "D",
    LactoDomDep == "Deplete" & SialidaseActivity == "High" & Leukocyte == "Low" ~ "E",
    LactoDomDep == "Dominant" & SialidaseActivity == "High" & Leukocyte == "Low" ~ "F",
    LactoDomDep == "Deplete" & SialidaseActivity == "Low" & Leukocyte == "Low" ~ "G",
    LactoDomDep == "Dominant" & SialidaseActivity == "Low" & Leukocyte == "Low" ~ "H"),
    BirthGestation_Weeks = BirthGestation/7) %>%
  modify_if(is.character, as.factor) %>%
  select(StatGroup, BirthGestation, BirthGestation_Weeks) %>%
  split(.$StatGroup) %>%
  map(summary)
```

***

## Table S6

The results from logistic regression analyses. The dependent variable was birth outcome (preterm/term) and independent variables were *Lactobacillus* abundance, sialidase activity and leukocyte wet mount results. 

> Data 

```{r}
# Prepare data for logistic regression
subcohort_lg <- subcohort %>%
  select(Group_PretermTerm, SialidaseActivity, Leukocyte, LactoDomDep) %>%
  mutate(SialidaseActivity = case_when(
    SialidaseActivity == "High" ~ 1,
    SialidaseActivity == "Low" ~ 0),
    Leukocyte = case_when(
      Leukocyte == "High" ~ 1,
      Leukocyte == "Low" ~ 0),
    Group_PretermTerm = case_when(
      Group_PretermTerm == "Preterm" ~ 1,
      Group_PretermTerm == "Term" ~ 0),
    LactoDomDep = case_when(
      LactoDomDep == "Deplete" ~ 1,
      LactoDomDep == "Dominant" ~ 0))

# Assign dependent variable as factor 
subcohort_lg[,1] <- as.factor(subcohort_lg[,1]) 
```

> Exploratory Analysis

**Boxplots**

```{r fig.width=9, fig.height=9}
# Aesthetics 
main_theme <- theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 15),
        axis.title.x = element_text(size = 15),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 14),
        strip.background = element_rect(color = "black", size = 1, linetype = "solid"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 14),
        legend.position = "bottom")

# Variables to loop 
colNames <- names(subcohort_lg)[c(2:4)]

# Boxplots (spread)
list_bp <- list()

for(i in colNames){
  bp <- ggplot(subcohort_lg, aes(x = .data[[i]], fill = subcohort_lg[,1])) + 
    stat_boxplot(geom = 'errorbar') +
    geom_boxplot() +
    scale_fill_manual(values = c("#ff595e", "#1982c4")) +
    labs(fill = "Group_PretermTerm") +
    theme_bw() +
    main_theme + 
    theme(axis.text.y = element_blank(),
          axis.ticks.y.left = element_blank())
  
  list_bp[[i]] <- bp
  }

wrap_plots(list_bp, ncol = 1)

# Descriptive stats 
subcohort_lg %>% 
  psych::describe()
```

**Histograms**

```{r fig.width=9, fig.height=9, warning=FALSE, message=FALSE}
# Histograms (distribution)
list_hist <- list()

for(i in colNames){
  hist <- ggplot(subcohort_lg, aes(x = .data[[i]], fill = subcohort_lg[,1])) +
    geom_histogram() +
    scale_fill_manual(values = c("#ff595e", "#1982c4")) +
    labs(fill = "sPTB1") + 
    theme_bw() +
    main_theme
  
  list_hist[[i]] <- hist
}

wrap_plots(list_hist, ncol = 1)
```

**Density Plots**

```{r fig.width=9, fig.height=9}
# Density plots (distribution)
list_dp <- list()

for(i in colNames){
  dp <- ggplot(subcohort_lg, aes(x = .data[[i]], color = subcohort_lg[,1])) + 
    geom_density(outline.type = "full", size = 0.7) +
    scale_color_manual(values = c("#168aad", "#ff4d6d")) + 
    labs(color = "sPTB1") +
    theme_bw() + 
    main_theme
  
  list_dp[[i]] <- dp
}

wrap_plots(list_dp, ncol = 2)
```

> Test Assumptions 

**(a) Appropriate Outcome Structure** 

YES, outcome binary as defined by preterm (1) or term (0) births.

**(b) Observation Independence** 

YES, none of the data points are replicates.

**(c) Large Sample Size (n 10 in each group)**

**Categorical Variables**

```{r}
# Count group sizes 
for(i in c(2:4)) {
  ct <- count(subcohort_lg, subcohort_lg[i])
  print(ct)
}
```

**(d) No Multi-collinearity**

- Correlation plot significance levels: 0, `***` (0.001), `**` (0.01), `*`, `.` (0.05), "` `" (0.1), 1
- Interpretation: `0 to 0.3` = small/weak, `0.3 to 0.5` = medium/moderate, `0.5 to 1.0` = large/strong

```{r fig.width=9, fig.height=9, warning=FALSE}
# Convert preterm/term group to numeric for correlation 
subcohort_ptb <- subcohort_lg %>% modify_if(is.factor, as.numeric)

# Correlation plots 
corr_subcohort_lg <- select(subcohort_ptb, 1:4) # only independent variables 
chart.Correlation(corr_subcohort_lg, method = "spearman", histogram = TRUE, pch = 19,
                  exact = FALSE)
mtext("Correlation Plot: Spearman Method (non-parametric)", line = 3)
```

**(e) Linearity Assumption** 

No continuous variables were used. 

> Logistic Regression & Model Diagnostics

**Regression**

```{r}
# Logistic Regression 
glm.fit1 <- glm(Group_PretermTerm ~ LactoDomDep + SialidaseActivity + Leukocyte, 
                data = subcohort_lg, family = binomial)

summary(glm.fit1)

# AIC 
print("Akaike Information Criteria (AIC)")
drop1(glm.fit1)

# Null and residual deviance
print("Null and residual deviance")
summary(glm.fit1)$deviance / summary(glm.fit1)$subcohort_lg.residual

# Wald test (https://www.r-bloggers.com/2015/08/evaluating-logistic-regression-models/)
regTermTest(glm.fit1, "LactoDomDep")
regTermTest(glm.fit1, "SialidaseActivity")
regTermTest(glm.fit1, "Leukocyte")

# Odds ratio (https://stackoverflow.com/questions/41384075/r-calculate-and-interpret-odds-ratio-in-logistic-regression)
exp(coef(glm.fit1))
```

> Diagnostics

**Influential Values (Outliers)**

```{r}
# Outliers 
outliers_1 <- check_outliers(glm.fit1)
plot(outliers_1, type = "dots")

# Cook's Distance 
plot(glm.fit1, which = 4, id.n = 5)

# Get data 
data.glmfit1 <- augment(glm.fit1) %>% mutate(index = 1:n()) # Extract model results 

# Top n largest values from Cook's distance 
print("Cook's Distance Influential Values")
data.glmfit1 %>% top_n(5, .cooksd) %>% reactable(rownames = TRUE)
```

**Residuals** 

```{r echo=TRUE}
# Binned Residuals 
binned_residuals(glm.fit1) 

# Standardized residuals 
ggplot(data.glmfit1, aes(index, .std.resid)) + 
  geom_point(aes(color = subcohort_lg[,1]), alpha = .5) +
  scale_color_manual(values = c("#168aad", "#ff4d6d")) + 
  labs(title = "Standardised Residuals", color = "sPTB1") +
  theme_bw() + 
  main_theme

# Potential influential data points with abs(.std.res) > 3 (i.e. outliers)
print("Influential Standardized Residuals")
data.glmfit1 %>% filter(abs(.std.resid) > 3) %>% reactable(rownames = TRUE)
```

**Multicollinearity**

```{r}
# Check multicollinearity using Variance Inflation Factor (VIF) - <5-10 is good 
print("Variance Inflation Factor (VIF) - <5-10 is good")
car::vif(glm.fit1)

# Multicollinearity 
multicol_1 <- check_collinearity(glm.fit1)
plot(multicol_1) +
  guides(x = guide_axis(angle = 90))
```

***

# Additional Analyses

## DNA Extraction Kit 

### Lactobacillus Abundance (Dom/Dep)

```{r warning=FALSE, fig.width=13, fig.height=5}
# Fisher's Exact Test
set.seed(1)
fisher.test(finalcohort$DNA_Extraction_Kit, finalcohort$LactoDomDep, 
            simulate.p.value = TRUE, B = 2000)

# Plots
domdep_colors <- c("#e31a1c", "#1f78b4")
func_barchart_stackedbar(data = finalcohort, 
                         x_var = finalcohort$DNA_Extraction_Kit, 
                         fill_var = finalcohort$LactoDomDep, 
                         fill_var_colors = domdep_colors, 
                         title = "DNA Extraction Kit x Lacto (Dom/Dep)", 
                         lab_x = "DNA_Extraction_Kit", 
                         lab_fill = "Lactobacillus",
                         x_guide = guide_axis())
```

### VMG

```{r warning=FALSE, fig.width=13, fig.height=6}
# Fisher's Exact Test
set.seed(1)
fisher.test(finalcohort$DNA_Extraction_Kit, finalcohort$VMG, 
            simulate.p.value = TRUE, B = 2000)

# Plots
func_barchart_stackedbar(data = finalcohort, 
                         x_var = finalcohort$DNA_Extraction_Kit, 
                         fill_var = finalcohort$VMG, 
                         fill_var_colors = vmg_color, 
                         title = "DNA Extraction Kit x VMG", 
                         lab_x = "DNA_Extraction_Kit", 
                         lab_fill = "VMG",
                         x_guide = guide_axis())
```

### Birth Outcome (Preterm/Term)

```{r warning=FALSE, fig.width=13, fig.height=6}
# Fisher's Exact Test
set.seed(1)
fisher.test(outcomecohort$DNA_Extraction_Kit, outcomecohort$Group_PretermTerm, 
            simulate.p.value = TRUE, B = 2000)

# Plots
ptb_colors <- c("#ff595e", "#1982c4")
func_barchart_stackedbar(data = outcomecohort, 
                         x_var = outcomecohort$DNA_Extraction_Kit, 
                         fill_var = outcomecohort$Group_PretermTerm, 
                         fill_var_colors = ptb_colors, 
                         title = "DNA Extraction Kit x Birth Outcome", 
                         lab_x = "DNA_Extraction_Kit", 
                         lab_fill = "Birth Outcome",
                         x_guide = guide_axis())
```

***

## Mode of Labour 

```{r}
outcomecohort_mol <- filter(outcomecohort, Mode_of_Labour != "Miscarriage")
```


### Lactobacillus Abundance

```{r warning=FALSE, fig.width=13, fig.height=6}
# Fisher's Exact Test
set.seed(1)
fisher.test(outcomecohort_mol$Mode_of_Labour, outcomecohort_mol$LactoDomDep, 
            simulate.p.value = TRUE, B = 2000)

# Plots
domdep_colors <- c("#e31a1c", "#1f78b4")
func_barchart_stackedbar(data = outcomecohort_mol, 
                         x_var = outcomecohort_mol$Mode_of_Labour, 
                         fill_var = outcomecohort_mol$LactoDomDep, 
                         fill_var_colors = domdep_colors, 
                         title = "Mode of Labour x Lacto (Dom/Dep)", 
                         lab_x = "Mode of Labour", 
                         lab_fill = "Lactobacillus",
                         x_guide = guide_axis(angle = 45))
```

### VMG

```{r warning=FALSE, fig.width=13, fig.height=6}
# Fisher's Exact Test
set.seed(1)
fisher.test(outcomecohort_mol$Mode_of_Labour, outcomecohort_mol$VMG, 
            simulate.p.value = TRUE, B = 2000)

# Plots
func_barchart_stackedbar(data = outcomecohort_mol, 
                         x_var = outcomecohort_mol$Mode_of_Labour, 
                         fill_var = outcomecohort_mol$VMG, 
                         fill_var_colors = vmg_color, 
                         title = "Mode of Labour x VMG", 
                         lab_x = "Mode of Labour", 
                         lab_fill = "VMG",
                         x_guide = guide_axis(angle = 45))
```

***

# Packages 

```{r}
sessionInfo()
```

